A comparative risk assesment of U.S. regional active transport using the ITHIM model

### 2016-10-05 R/ITHIM.Rmd
Edits to manuscript.

#geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+")) +
theme_tufte(base_size=12, ticks=T)
#

# + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))
# + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
```

>>>>

#geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
scale_x_discrete(name="", labels = c("5-14","15-29","30-44","45-59","60-69","70-79","80+"))
# + theme_tufte(base_size=12, ticks=T)
#
# + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))
# + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
```

### 2016-10-05 manuscript/manuscript.tex
Edits to manuscript.

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\ithim}{\emph{ITHIM}}
\newcommand{\R}{\emph{R}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}
\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
\newcommand{\af}{\textrm{PAF}}
\newcommand{\mets}{MET-hrs./week}
\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
\newcommand{\vmt}{$\mathrm{VMT}$}

@@

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of travel-related exposure and non-travel-related
exposure, which are modeled as two independent random variables.  The
age-sex specific mean ratios for non-travel-related exposure are
estimated from the ATUS \cite{ATUS} and the referent group mean is
treated as a variable.  We see in Figure \ref{dalyFigure} that varying
it does not greatly affect the estimate for disease burden.

\subsection{Active Transport}

\textit{The time spent walking or cycling, i.e., active transport
  time, is modeled as a log-normal random variable with means defined
  for each age-sex stratum.  The non-travel-related physical activity
  is also modeled as log-normal with stratum-specific means.}

\textit{ The travel-related exposure is estimated using active
  transport time along with assumptions about how much physical
  activity is required for cycling and walking.  Active transport time
  and non-travel-related physical activity are modeled as independent
  random variables.  A fixed value was used as the coefficient of
  variation for the distributin of active transport time. A different,
  but also fixed, value was used for the non-travel-related physical
  activity distribution.}

@@

%% \textit{ Can we model the decrease in \pEm{2.5} nationally as a
%%   function of decrease in driving time?  The original plan was to use
%%   a baseline value for \pEm{2.5} and a linear relationship bewteen
%%   change in \pEm{2.5} and change in \vmt{}.}

%% \begin{figure}[t]
%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
%%     \caption{Counts}\label{counts}
%% \end{figure}

\section{New York and California}

\begin{itemize}
\item We used the National Household Transportation Survey \cite{NHTS} to
  estimate the mean walk and cycle times for individuals that live in
  metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
  corresponds to individuals whose homes are in an MSA or cMSA)
  \bi\item Data quality issues along with assumptions about travel speed
created mirky data.  The median values of trip time were used to avoid
undue influence by outliers.\ei
\item We then divided these into the 10 regions of the U.S. defined by
  the Department of Health and Human Services (HH1-10)
\item Using only the active transport component of the ITHIM model we
  ran comparative analyses for each of the ten regions, using the
  national averages as the baseline.
\item We used data from the American Time Use Survey \cite{ATUS} to
  estimate the relative means for each of the age-sex strata.  We then
  treat the mean of the referent group as a variable and compute
  results over a grid of values for this variable.  See Figure
  \ref{dalyFigure}.
\item Values for DALYs were found using the CDC Wonder database \cite{CDCWonder}
\end{itemize}

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
    \caption{Mean travel time by mode (walking/cycling) for each of
      the four regions with a decrease in DALYs
      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
      national means are displayed on the far right.  These four
      regions correspond to New England/NewYork and the West
      Coast. The age classes are defined as in the original ITHIM
      implementation using 8 eight age classes defined in terms of
      years as follows: 1. 0--4, 2. 5--14, 3. 15--29, 4. 30--44,
      5. 45--59, 6. 60--69, 7. 70--79, and 8. 80+. Age class 1 is
      omitted from all calclulations in this paper due to lack of
      data.}\label{meanMatrices}
\end{figure}

%\bibliographystyle{natbib}
%\bibliographystyle{bmc_article}
\bibliographystyle{mypapers}
\bibliography{ITHIM}
\end{document}

>>>>

\newcommand{\col}[2][red]{\textcolor{#1}{#2}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\package}{\emph{ITHIM}}
\newcommand{\ithim}{\emph{ITHIM}}
\newcommand{\R}{\emph{R}}
\newcommand{\sgy}[1]{{\itshape\col{#1}}}
\newcommand{\XX}{\sgy{XX}}
\newcommand{\webpage}{\url{http://www.ithim.ghi.wisc.edu}}
\newcommand{\af}{\textrm{PAF}}
\newcommand{\mets}{MET-hrs./week}
\newcommand{\pEm}[1]{$\mathrm{PM}_{#1}$}
\newcommand{\vmt}{$\mathrm{VMT}$}
\newcommand{\logNormal}{log-normal}

@@

Here $R(x)$ represents the risk when the individual has exposure $x$,
physical activity (\mets).  $P(x)$ and $Q(x)$ denote the population
density of individuals with exposure $x$, in the baseline and
alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of two independent random variables, travel-related
exposure and non-travel-related exposure.  Non-travel-related exposure
is modeled with a \logNormal{} distribution with the age-sex specific
mean ratios for non-travel-related exposure are estimated from the
ATUS \cite{ATUS}.  The referent group mean is treated as a variable
and the coefficient of variation is treated as constant across age-sex
strata and estimated with ATUS.  We see in Figure \ref{dalyFigure}
that varying the mean values for non-travel-related exposure does not
greatly affect the estimate for disease burden.  The distribution for
travel-related exposure is estimated from the time spent walking or
cycling, i.e., active transport time, which as in the original ITHIM
model has a \logNormal{} distribution with constant coefficient of
variation across age-sex strata \cite{}.  Travel-related exposure
is estimated from active transport time using assumptions about how
much physical activity is required for cycling and walking.

\section{Results}

We used the National Household Transportation Survey \cite{NHTS} to
estimate the mean walk and cycle times for individuals that live in
metropolitan areas for each of seven states (CA, TX, NY, FL, VA, NC
and AZ) which had sufficient sample size for reasonable estimates of
age-sex specific travel-related walking and cycling means, Figure
\ref{meanMatrices}.  Of these seven states two were found to have
active transport behavior that would avert thousands of DALYs if
adopted nation-wide. Values for national disease-specific DALYs were
found using the CDC Wonder database \cite{CDCWonder}.

%% \begin{itemize}
%% \item We used the National Household Transportation Survey \cite{NHTS} to
%%   estimate the mean walk and cycle times for individuals that live in
%%   metropolitan areas ($\textrm{DAYV2PUB.MSA} \in (1,2,\ldots,5)$ which
%%   corresponds to individuals whose homes are in an MSA or cMSA)
%%   \bi\item Data quality issues along with assumptions about travel speed
%% created mirky data.  The median values of trip time were used to avoid
%% undue influence by outliers.\ei
%% \item We then divided these into the 10 regions of the U.S. defined by
%%   the Department of Health and Human Services (HH1-10)
%% \item Using only the active transport component of the ITHIM model we
%%   ran comparative analyses for each of the ten regions, using the
%%   national averages as the baseline.
%% \item We used data from the American Time Use Survey \cite{ATUS} to
%%   estimate the relative means for each of the age-sex strata.  We then
%%   treat the mean of the referent group as a variable and compute
%%   results over a grid of values for this variable.  See Figure
%%   \ref{dalyFigure}.
%% \item 
%% \end{itemize}

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
    \caption{Mean travel time by mode (walking/cycling) for each of
      seven states National means are displayed on the right.}\label{meanMatrices}
\end{figure}

%\bibliographystyle{natbib}
\bibliographystyle{bmc_article}
%\bibliographystyle{mypapers}
\bibliography{ITHIM}
\end{document}

@@

%% \textit{ Can we model the decrease in \pEm{2.5} nationally as a
%%   function of decrease in driving time?  The original plan was to use
%%   a baseline value for \pEm{2.5} and a linear relationship bewteen
%%   change in \pEm{2.5} and change in \vmt{}.}
%% \begin{figure}[t]
%%     \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
%%     \caption{Counts}\label{counts}
%% \end{figure}
